Embedding Intuitionistic into Classical Logic

### 2023-03-19 LPAR/LPAR-EIICL.tex
Formatting
Added newpage
typos

Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We now given an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

	On the other hand consider \[\forall w\phi = \forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))\in\mathcal X^{bS}\] and $\psi_1\dots\psi_n\in M_T$. Let $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)$. If $\mathcal M, v[m/w]\models C(w)$ it follows that $\mathcal M_T, v[\psi_1\dots\psi_m]\models C(w)$ and we are done. Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$. The former is impossible since $g_\psi$ is increasing for all $\psi$. For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$, $\mathcal M, v[m/w]\models A(f_\psi(w))$ and $\mathcal M, v[m/w]\not\models B(f_\psi(w))$.

	There are now 2 cases. Suppose $\psi$ is fulfilled at $m$. Then we have $\mathcal M, v[m/w]\not\models A(w)\to B(w)$ and $g_\psi(m) = m$ and therefore $\mathcal M, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f\psi(w))$, i.e. $\mathcal M, v[\psi_1\dots\psi_n/w]\models \phi$.

	On the other hand suppose $\psi$ is not fulfilled at $m$. Then $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and $g_\psi(m) = f^I_\psi(m)$. In this case it directly follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and thus $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e. $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.

>>>>

Suppose we are given a formula $\varphi$ in normal form $$\left(\bigwedge\mathcal R\wedge\bigwedge\mathcal X\right)\to P$$ as in Lemma~\ref{lemma:propositional-normal-form}. We now give an intuition on how to obtain a QBF of polynomial size that expresses that $\varphi$ has an intuitionistic counter-model:
Instead of expressing validity at every node of $\mathcal M_T$ --- which was done in the CPC translation --- we express that on each path in $\mathcal M_T$ the nodes satisfy all conditions for $\mathcal M_T$ being a counter-model.
The universally quantified variables $X_\psi$ in the next definition handle the branching, i.e. express which path is considered.

@@

	On the other hand consider \[\forall w\phi = \forall w(s\preceq w\to(w\preceq f_\psi(w)\to A(f_\psi(w))\to B(f_\psi(w)))\to C(w))\in\mathcal X^{bS}\] and $\psi_1\dots\psi_n\in M_T$. Let $m = g_{\psi_n}(\dots(g_{\psi_1}(s^I))\dots)$. If $\mathcal M, v[m/w]\models C(w)$ it follows directly that $\mathcal M_T, v[\psi_1\dots\psi_m]\models C(w)$ and we are done. Otherwise $\mathcal M, v[m/w]\not\models s\preceq w$ or $\mathcal M, v[m/w]\not\models w\preceq f_\psi(w)\to A(f_\psi(w)) \to B(f_\psi(w))$. The former is impossible since $g_\psi$ is increasing for all $\psi$. For the latter to hold we must have $\mathcal M, v[m/w]\models w\preceq f_\psi(w)$ as well as $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$. Since by definition $\mathcal M_T, v[\psi_1\dots\psi_m/w]\models s\preceq w\wedge w\preceq f_\psi(w)$ we must show $\mathcal M_T, v[\psi_1\dots\psi_m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$.

	There are now 2 cases. Suppose $\psi$ is fulfilled at $m$. Then we have $\mathcal M, v[m/w]\not\models A(w)\to B(w)$ and $g_\psi(m) = m$ and therefore $\mathcal M, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e. indeed $\mathcal M, v[\psi_1\dots\psi_n/w]\models\phi$.

	On the other hand suppose $\psi$ is not fulfilled at $m$, i.e. $g_\psi(m) = f^I_\psi(m)$. Since $\mathcal M, v[m/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ it directly follows that $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$ and thus $\mathcal M_T, v[\psi_1\dots\psi_n/w]\not\models A(f_\psi(w))\to B(f_\psi(w))$, i.e. indeed $\mathcal M_T, v[\psi_1\dots\psi_n/w]\models\phi$.
